Sparse Function-space Representation of Neural Networks

### 2023-08-31 workshops/icml-duality/main.tex
updated

\begin{abstract}
Deep neural networks are known to lack uncertainty estimates, struggle to incorporate new data, and suffer from catastrophic forgetting. We present a method that mitigates these issues by converting neural networks from weight-space to a low-rank function-space representation, via the so-called dual parameters. In contrast to previous work, our sparse representation captures the joint distribution over the entire data set, rather than only over a subset. This offers a compact and principled way of capturing uncertainty and enables us to incorporate new data without retraining whilst retaining predictive performance. We provide proof-of-concept demonstrations with the proposed approach for quantifying uncertainty in supervised learning on UCI benchmark tasks.
\end{abstract}

\section{Introduction}
%
Deep learning \citep{goodfellow2016deep} has become the cornerstone of contemporary artificial intelligence, proving remarkably effective in tackling supervised and unsupervised learning tasks in the {\em large data}, {\em offline}, and {\em gradient-based training} regime. Despite its success, gradient-based learning techniques exhibit limitations. Firstly, how can we efficiently quantify uncertainty without resorting to expensive and hard-to-interpret sampling in the model's weight-space? Secondly, how to update the weights of an already trained model with new batches of data without compromising the performance on past data? These questions become central when applied to sequential learning paradigms, such as continual learning \citep[CL,][]{parisi2019continual, de2021continual} and reinforcement learning  \citep[RL,][]{sutton2018reinforcement}. In CL, access to the previous data is lost, and then the challenge is retaining a compact representation of the problem to alleviate forgetting over the life-long learning horizon~\citep{mccloskey1989catastrophic}. Similarly, in RL, the model must adapt to environmental observations through exploration, while leveraging prediction uncertainties to assess potential future paths.\looseness-1

\begin{figure}[t!]
  % \centering\scriptsize
  \centering\Large
  % Figure options
  \pgfplotsset{axis on top,scale only axis,width=\figurewidth,height=\figureheight, ylabel near ticks,ylabel style={yshift=-2pt},y tick label style={rotate=90},legend style={nodes={scale=1., transform shape}},tick label style={font=\Large,scale=1}}
  \pgfplotsset{xlabel={Input, $x$},axis line style={rounded corners=2pt}}
  % Set figure
  \setlength{\figurewidth}{.4\textwidth}
  \setlength{\figureheight}{\figurewidth}
  %
  \def\inducing{\Huge Sparse inducing points}
  %
  \begin{subfigure}[c]{.52\columnwidth}
    \raggedleft
    \pgfplotsset{ylabel={Output, $y$}}
    \input{./fig/regression-nn.tex}%
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{.02\columnwidth}
    \centering
    \tikz[overlay,remember picture]\node(p0){};
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{.4\columnwidth}
    \raggedleft
    \pgfplotsset{yticklabels={,,},ytick={\empty}}
    \input{./fig/regression-nn2svgp.tex}%
  \end{subfigure}
%  \hfill
%  \begin{subfigure}[c]{.01\textwidth}
%    \centering
%    \tikz[overlay,remember picture]\node(p1){};
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[c]{.28\textwidth}
%    \raggedleft
%    \pgfplotsset{yticklabels={,,},ytick={\empty}}
%    \input{./fig/regression-update.tex}%
%  \end{subfigure}
  % \vspace*{-1em}
  \caption{\textbf{Regression example on an MLP with two hidden layers.} Left:~Predictions from the trained neural network. Right:~Our approach summarizes all the training data with the help of a set of inducing points.} %The model captures the predictive mean and uncertainty, and (right) makes it possible to incorporate new data without retraining the model.}
  \label{fig:teaser}
  %
  \begin{tikzpicture}[remember picture,overlay]
    % Arrow style
    \tikzstyle{myarrow} = [draw=black!80, single arrow, minimum height=14mm, minimum width=2pt, single arrow head extend=4pt, fill=black!80, anchor=center, rotate=0, inner sep=5pt, rounded corners=1pt]
    % Arrows
    \node[myarrow] (p0-arr) at ($(p0) + (0.2em,1.2em)$) {};
    % \node[myarrow] (p0-arr) at ($(p0) + (1em,1.5em)$) {};
    %\node[myarrow] (p1-arr) at ($(p1) + (1em,1.5em)$) {};
    % Arrow labels
    \node[font=\scriptsize\sc,color=white] at (p0-arr) {\our};
    %\node[font=\scriptsize\sc,color=white] at (p1-arr) {new data};
  \end{tikzpicture}
  \vspace*{-1em}
\end{figure}

@@

% by linearising a neural network around the Maximum a Posteriori (MAP) weights.
% a GP posterior -- can be obtained
%
In supervised learning, given a data set $\dataset = \{(\vx_{i} , \vy_{i})\}_{i=1}^{N}$, with input $\vx_i \in \inputDomain$ and output $\vy_i \in \outputDomain$ pairs, the weights $\weights \in \R^{P}$ of a neural network, $f_\mathbf{w}: \inputDomain \to \outputDomain$ (yet, to simplify the notation we restrict presentation to scalar output), are usually trained to minimize the (regularized) empirical risk,\looseness-1
%
\begin{align} \label{eq-empirical-risk}
  \weights^{*} &=
  \arg \min_{\weights} \mathcal{L}(\dataset,\weights) \nonumber \\ &=
  \arg \min_{\weights} \textstyle\sum_{i=1}^{N} \ell(f_\weights(\mathbf{x}_{i}), y_i) + \delta \mathcal{R}(\weights).
\end{align}
%
If $\ell(f_\weights(\vx_{i}), y_i) = -\log(p(y \mid f_\weights(\vx_{i}))$ and the regularizer $\mathcal{R}(\weights) = \frac{1}{2}\|\weights\|^{2}_2$, then we can view \cref{eq-empirical-risk} as the maximum {\it a~posteriori} (MAP) solution to a Bayesian objective, where the regularization weight takes the role of a prior precision parameter, \ie, $p(\vw) = \Norm(\vzeros, \delta^{-1} \MI)$.
%Bayesian inference offers a principled approach to quantifying uncertainty in neural networks.
%The goal is to find the posterior over the weights ${p(\vw \mid \vy) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ as it represents our belief in the parameters after combining data $\dataset$ with our prior $p(\vw)$.
%Although the true posterior $p(\vw \mid \dataset)$ is intractable given the non-linearities of the neural network, one can resort to sampling techniques such as Hamiltonian Monte-Carlo (HMC).
%However, for most applications their high computational costs make them impractical so we need approximate inference techniques.
%In \cref{fig:banana}, we show a qualitative example of the induced function-space posterior obtained with our method compared to the posterior obtained through HMC sampling on the Banana toy dataset.
The posterior over the weights ${p(\vw \mid \dataset) \propto p(y \mid f_{\weights}(\vx)) \, p(\weights)}$ is generally not available in closed form. Sampling methods that characterize the posterior with a finite set of samples in weight-space are general-purpose, but computationally heavy and impractical for downstream applications.

\textbf{Neural network function-space} %\todo{Could be titled some other way, like Function-space representation}
%As a neural network is a deterministic mapping, the weight-space posterior induces a distribution over the function values.
%Intuitively, if one was to sample from the weight posterior, the corresponding functions created can
%be viewed as perturbed versions of the function at the MAP estimate $f_{\vw^*}$.
%In most applications, we care about predictions from the neural network and not the weights themselves.
%As such, it is the distribution over function values that we are actually interested in.
Neural networks are deterministic parametric functions, but even if the training is typically an optimization in the weight-space, in most applications, we are interested in the parametrized function and not in the weights themselves.
%While neural networks are deterministic mappings defined by their weights, the ultimate goal of training a neural network is to optimize the function it represents, not the weights themselves.
Consequently, the weight-space posterior corresponds to a distribution over function values.
Intuitively, if one was to sample from the weight posterior, the corresponding functions created can be viewed as perturbed versions of the function at the MAP estimate $f_{\vw^*}$. This perspective aligns better with the main objective of making representative predictions given the observations.

\textbf{Linearization gives rise to a Gaussian process}
Our goal is to capture the distribution over the neural network model functions through their first two moments. The first two moments characterize a Gaussian process with a mean function $\mu(\cdot)$ and a covariance function (kernel) $\kappa(\cdot,\cdot)$.
Recent work \citep{khan2019approximate,maddox2021fast} has shown that linearizing approximations in the weight space lead to function-space equivalent approximations.
As Gaussian distributions remain tractable under linear transformations, a linear function in terms of parameters can be converted from the weight space to the function space \citep[see Ch.~2.1 in ][]{rasmussen2006gaussian} as follows:
%
\begin{align} \label{eq:weight_func}
 &f_\weights(\vx) \approx
%g_\weights(\mathbf{x}) =
 \phi^\top\!(\vx) \, \vw \quad\implies \nonumber \\
 &\mu(\vx) = 0 \quad \text{and} \quad \kappa(\vx, \vx') = \frac{1}{\delta} \phi^\T\!(\vx) \, \phi(\vx').
\end{align}
% The posterior structure directly relates to the optimization loss around the MAP weights $\vw^*$.
A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN~\citep{khan2019approximate, daxberger2021laplace, maddox2021fast}.
The Laplace-GGN takes the MAP solution and approximates the Hessian of the loss function
with the GGN.
Following the work of \citet{khan2019approximate}, we can view this as a linear approximation of the neural network, as $f_{\weights^*}(\vx) \approx
%g_{\weights}(\vx) =
\Jac{\weights^*}{\vx} \, \weights$, where $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{1 \times P}$ is the Jacobian at the MAP.
For notational conciseness, we % \cref{eq-laplace-approx-function-space},
restrict our notatin to a single function output. The extension to multiple outputs is straightforward.

Using the Hessian of the approximate model, we arrive at the Laplace-GGN approximate posterior over $\vw$.
Therefore, the Laplace-GGN linear model can be used to convert our weight space model to the function space,
\begin{equation}
\label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx) =  0 \quad \text{and} \quad
  \kappa(\vx, \vx')
  = \frac{1}{\delta} \Jac{\weights^*}{\vx} \, \JacT{\weights^*}{\vx'},
\end{equation}
where the kernel is the so-called Neural Tangent Kernel \citep[NTK,][]{jacot2018neural}.
We can combine this kernel function with the data set $\dataset$ to construct the posterior.
%
\citet{khan2019approximate} took a similar approach, but instead of fitting the GP posterior to the actual $\vy$, they rely on a transformation of $\vy$. Similarly, \citet{immer2021improving} obtain the same covariance function but use a different mean because they rely on a first-order approximation of the neural network to form a Bayesian GLM model. Both approaches adjust the GP posterior mean function to ensure the predictions are from the neural network $f_{\vw^*}(\vx)$.
In contrast, our method makes predictions directly from the GP and avoids such adjustments.
In the next section, we present our method for converting a trained NN to a sparse functional representation.
% on the non-sparse approximation.
% of our
% based on the non-sparse approximation.
% we will present the general formulation of the sparse functional representation of our trained neural network based on the non-sparse approximation.

\newpage
\section{\our: Sparse function-space representation of neural networks}\label{sec:sfr}
%\paragraph{GP in the dual parameters}
\todo{update this with paul}
The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
%
\begin{align}  \label{eq:gp_pred}
  \myexpect_{p(f_i \mid\vy)}[f_i] &= \vk_{\vx i}^\top \valpha, \\
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i}, \nonumber
\end{align}
%
where $f_\vw(\vx_i) = f_i$ is the function output at input $\vx_i$
and the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.
\cref{eq:gp_pred} states that the first two moments of the resultant posterior process (which may not be a GP), can be parameterized via the dual
parameters $\valpha, \vbeta \in \R^{N}$,
% which are the vectors
defined as, \looseness-1
%
\begin{equation}
\begin{split}
  \label{eq:dual_param}
  \alpha_i &\coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f}\log p(y_i \mid f) |_{f=f_i}], \\
  %\quad \text{and} \quad
  \beta_i &\coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f f}\log p(y_i \mid f_i) |_{f=f_i}].
  \end{split}
\end{equation}
%
The relationships specified are valid for generic likelihoods, and involve no approximations since the expectation is under the exact posterior, but given that the model can be expressed in a kernel formulation. \cref{eq:gp_pred} and \cref{eq:dual_param} highlight that the approximate inference technique, usually viewed as a posterior approximation, can be alternatively interpreted as an approximation of the expectation of gradients of loss/likelihoods.

Originally \citet{csato2002sparse} iteratively find dual variables using an expectation propagation \citep[EP,][]{minka2001expectation} method. More recently, \citet{khan2017conjugate,adam2021dual} have shown this relationship for variational Gaussian processes where the above expectation is with respect to the approximate variational posterior. Meanwhile in \citet{wilkinson2023bayes}, they show links between linearization methods and how they solve the \cref{eq:dual_param}.

\paragraph{Dual parameters from NN}
Given that we use a Laplace approximation of the neural network, we remove the expectation over the posterior \citep[see Ch.~3.4.1 in][for derivation]{rasmussen2006gaussian} and we get the following formulation of the dual variables,
%
\begin{align}
  \label{eq:dual_param_laplace}
  \hat{\alpha}_i &\coloneqq \nabla_{f}\log p(y_i \mid f) |_{f=f_i} , \\
  %\quad \text{and} \quad
  \hat{\beta}_i &\coloneqq - \nabla^2_{ff}\log p(y_i \mid f) |_{f=f_i}.
\end{align}
%
Substituting \cref{eq:dual_param_laplace} into \cref{eq:gp_pred}, we obtain our GP model based on the converged neural network. Again, this is similar to what was derived in \citet{immer2021improving} for the posterior variance function. However, they use the $f_{\vw^*}$ for the posterior mean and do not emphasize the significance of the dual variables. The problem with \cref{eq:gp_pred} is that to make predictions and compute variances we must incur a cost of $O(N^3)$ which limits the use of the GP on large data sets.

\begin{table*}[t!]
  \centering\scriptsize
  \caption{Comparisons and ablations on UCI data with negative log predictive density (NLPD\textcolor{gray}{\footnotesize$\pm$std}, lower better). Our sparse \our (with $M=20\% \text{ of } N$) is on par with the Laplace methods (BNN/GLM) and outperforms the GP subset method.} %Results for methods marked with * as reported in the original benchmark~\citep{immer2021improving}.} %See \cref{app:uci} for additional tables with comparisons.}
	\label{tbl:uci}
	% \vspace*{-4pt}

	% Control table spacing
	% \renewcommand{\arraystretch}{1.}
	% \setlength{\tabcolsep}{1.2pt}
	% \setlength{\tblw}{0.083\textwidth}

@@

\paragraph{Sparsifying the NN GP}
\label{sec:sparse-dual-gp}
%
Sparse Gaussian processes reduce the computational complexity by representing the GP as a low-rank approximation induced by a sparse set of input points \citep[see][for an early overview]{quinonero2005unifying}. Given that we have computed the dual variables derived from our neural network predictions and a kernel function, we could essentially employ any of these sparsification methods. In this work, we opt for the approach suggested by \citet{titsias2009variational} \citep[also used in the DTC approximation, see][]{quinonero2005unifying}, which defines the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) \, q(\vu) \, \mathrm{d}\vu$. Given that we have $p(f_i \mid \vu)$ determined by our GP prior, the goal is to find a $q(\vu)$.

>>>>

\begin{abstract}
Deep neural networks (NNs) are known to lack uncertainty estimates, struggle to incorporate new data, and suffer from catastrophic forgetting. We present a method that mitigates these issues by converting NNs from weight-space to a low-rank function-space representation, via the so-called dual parameters. In contrast to previous work, our sparse representation captures the joint distribution over the entire data set, rather than only over a subset. This offers a compact and principled way of capturing uncertainty and enables us to incorporate new data without retraining whilst retaining predictive performance. We provide proof-of-concept demonstrations with the proposed approach for quantifying uncertainty in supervised learning on UCI benchmark tasks.
\end{abstract}

\section{Introduction}
%
Deep learning \citep{goodfellow2016deep} has become the cornerstone of contemporary artificial intelligence, proving remarkably effective in tackling supervised and unsupervised learning tasks in the {\em large data}, {\em offline}, and {\em gradient-based training} regime. Despite its success, gradient-based learning techniques exhibit limitations. Firstly, how can we efficiently quantify uncertainty without resorting to expensive and hard-to-interpret sampling in the model's weight-space? Secondly, how to update the weights of an already trained model with new batches of data without compromising the performance on past data? These questions become central when applied to sequential learning paradigms, such as continual learning \citep[CL,][]{parisi2019continual, de2021continual} and reinforcement learning  \citep[RL,][]{sutton2018reinforcement}. In CL, access to the previous data is lost, and then the challenge is retaining a compact representation of the problem to alleviate forgetting over the life-long learning horizon~\citep{mccloskey1989catastrophic}.
Similarly, in RL, the model must adapt to environmental observations and can leverage predictive uncertainties for exploration.\looseness-1

\begin{figure}[t!]
  % \centering\scriptsize
  \centering\Large
  % Figure options
  \pgfplotsset{axis on top,scale only axis,width=\figurewidth,height=\figureheight, ylabel near ticks,ylabel style={yshift=-2pt},y tick label style={rotate=90},legend style={nodes={scale=1., transform shape}},tick label style={font=\Large,scale=1}}
  \pgfplotsset{xlabel={Input, $x$},axis line style={rounded corners=2pt}}
  % Set figure
  \setlength{\figurewidth}{.41\textwidth}
  \setlength{\figureheight}{\figurewidth}
  %
  \def\inducing{\Huge Sparse inducing points}
  %
  \begin{subfigure}[c]{.52\columnwidth}
    \raggedleft
    \pgfplotsset{ylabel={Output, $y$}}
    \input{./fig/regression-nn.tex}%
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{.02\columnwidth}
    \centering
    \tikz[overlay,remember picture]\node(p0){};
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{.4\columnwidth}
    \raggedleft
    \pgfplotsset{yticklabels={,,},ytick={\empty}}
    \input{./fig/regression-nn2svgp.tex}%
  \end{subfigure}
%  \hfill
%  \begin{subfigure}[c]{.01\textwidth}
%    \centering
%    \tikz[overlay,remember picture]\node(p1){};
%  \end{subfigure}
%  \hfill
%  \begin{subfigure}[c]{.28\textwidth}
%    \raggedleft
%    \pgfplotsset{yticklabels={,,},ytick={\empty}}
%    \input{./fig/regression-update.tex}%
%  \end{subfigure}
  \vspace*{-0.8em}
  \caption{\textbf{Regression example on an MLP with two hidden layers.} Left:~Predictions from the trained neural network. Right:~Our approach (\our) equips trained NNs with uncertainty estimates.} %The model captures the predictive mean and uncertainty, and (right) makes it possible to incorporate new data without retraining the model.}
  % \caption{\textbf{Regression example on an MLP with two hidden layers.} Left:~Predictions from the trained neural network. Right:~Our approach (\our) summarizes all the training data with the help of a set of inducing points.} %The model captures the predictive mean and uncertainty, and (right) makes it possible to incorporate new data without retraining the model.}
  \label{fig:teaser}
  %
  \begin{tikzpicture}[remember picture,overlay]
    % Arrow style
    \tikzstyle{myarrow} = [draw=black!80, single arrow, minimum height=14mm, minimum width=2pt, single arrow head extend=4pt, fill=black!80, anchor=center, rotate=0, inner sep=5pt, rounded corners=1pt]
    % Arrows
    \node[myarrow] (p0-arr) at ($(p0) + (0.2em,1.2em)$) {};
    % \node[myarrow] (p0-arr) at ($(p0) + (1em,1.5em)$) {};
    %\node[myarrow] (p1-arr) at ($(p1) + (1em,1.5em)$) {};
    % Arrow labels
    \node[font=\scriptsize\sc,color=white] at (p0-arr) {\our};
    %\node[font=\scriptsize\sc,color=white] at (p1-arr) {new data};
  \end{tikzpicture}
  \vspace*{-1em}
\end{figure}

@@

% by linearising a neural network around the Maximum a Posteriori (MAP) weights.
% a GP posterior -- can be obtained
%
In supervised learning, given a data set $\dataset = \{(\vx_{i} , \vy_{i})\}_{i=1}^{N}$, with input $\vx_i \in \inputDomain$ and output $\vy_i \in \outputDomain$ pairs, the weights $\weights \in \R^{P}$ of a NN, $f_\mathbf{w}: \inputDomain \to \outputDomain$ (yet, to simplify the notation we restrict presentation to scalar output), are usually trained to minimize the (regularized) empirical risk,\looseness-1
%
\begin{align} \label{eq-empirical-risk}
  \weights^{*} &=
  \arg \min_{\weights} \mathcal{L}(\dataset,\weights) \nonumber \\ &=
  \arg \min_{\weights} \textstyle\sum_{i=1}^{N} \ell(f_\weights(\mathbf{x}_{i}), y_i) + \delta \mathcal{R}(\weights).
\end{align}
%
If $\ell(f_\weights(\vx_{i}), y_i) = -\log(p(y \mid f_\weights(\vx_{i}))$ and the regularizer $\mathcal{R}(\weights) = \frac{1}{2}\|\weights\|^{2}_2$, then we can view \cref{eq-empirical-risk} as the maximum {\it a~posteriori} (MAP) solution to a Bayesian objective, where the regularization weight takes the role of a prior precision parameter, \ie, $p(\vw) = \Norm(\vzeros, \delta^{-1} \MI)$.
%Bayesian inference offers a principled approach to quantifying uncertainty in neural networks.
%The goal is to find the posterior over the weights ${p(\vw \mid \vy) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ as it represents our belief in the parameters after combining data $\dataset$ with our prior $p(\vw)$.
%Although the true posterior $p(\vw \mid \dataset)$ is intractable given the non-linearities of the neural network, one can resort to sampling techniques such as Hamiltonian Monte-Carlo (HMC).
%However, for most applications their high computational costs make them impractical so we need approximate inference techniques.
%In \cref{fig:banana}, we show a qualitative example of the induced function-space posterior obtained with our method compared to the posterior obtained through HMC sampling on the Banana toy dataset.
The posterior over the weights ${p(\vw \mid \dataset) \propto p(y \mid f_{\weights}(\vx)) \, p(\weights)}$ is generally not available in closed form. Sampling methods that characterize the posterior with a finite set of samples in weight-space are general-purpose, but computationally heavy and impractical for downstream applications.

\textbf{Neural network function-space} %\todo{Could be titled some other way, like Function-space representation}
%As a neural network is a deterministic mapping, the weight-space posterior induces a distribution over the function values.
%Intuitively, if one was to sample from the weight posterior, the corresponding functions created can
%be viewed as perturbed versions of the function at the MAP estimate $f_{\vw^*}$.
%In most applications, we care about predictions from the neural network and not the weights themselves.
%As such, it is the distribution over function values that we are actually interested in.
NNs are deterministic parametric functions, but even if the training is typically an optimization in the weight-space, in most applications, we are interested in the parametrized function and not in the weights themselves.
%While neural networks are deterministic mappings defined by their weights, the ultimate goal of training a neural network is to optimize the function it represents, not the weights themselves.
Consequently, the weight-space posterior corresponds to a distribution over function values.
Intuitively, if one was to sample from the weight posterior, the corresponding functions created can be viewed as perturbed versions of the function at the MAP estimate $f_{\vw^*}$. This perspective aligns better with the main objective of making representative predictions given the observations.

\textbf{Linearization gives rise to a Gaussian process}
Our goal is to capture the distribution over the NN model functions through their first two moments. The first two moments characterize a GP with a mean function $\mu(\cdot)$ and a covariance function (kernel) $\kappa(\cdot,\cdot)$.
Recent work \citep{khan2019approximate,maddox2021fast} has shown that linearizing approximations in the weight space lead to function-space equivalent approximations.
As Gaussian distributions remain tractable under linear transformations, a linear function in terms of parameters can be converted from the weight space to the function space \citep[see Ch.~2.1 in ][]{rasmussen2006gaussian} as follows:
%
\begin{align} \label{eq:weight_func}
 &f_\weights(\vx) \approx
%g_\weights(\mathbf{x}) =
 \phi^\top\!(\vx) \, \vw \quad\implies \nonumber \\
 &\mu(\vx) = 0 \quad \text{and} \quad \kappa(\vx, \vx') = \frac{1}{\delta} \phi^\T\!(\vx) \, \phi(\vx').
\end{align}
% The posterior structure directly relates to the optimization loss around the MAP weights $\vw^*$.
A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN~\citep{khan2019approximate, daxberger2021laplace, maddox2021fast}.
The Laplace-GGN takes the MAP solution and approximates the Hessian of the loss function
with the GGN.
Following the work of \citet{khan2019approximate}, we can view this as a linear approximation of the NN, as $f_{\weights^*}(\vx) \approx
%g_{\weights}(\vx) =
\Jac{\weights^*}{\vx} \, \weights$, where $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{1 \times P}$ is the Jacobian at the MAP.
For notational conciseness, we % \cref{eq-laplace-approx-function-space},
restrict our notatin to a single function output. The extension to multiple outputs is straightforward.

Using the Hessian of the approximate model, we arrive at the Laplace-GGN approximate posterior over $\vw$.
Therefore, the Laplace-GGN linear model can be used to convert our weight space model to the function space,
\begin{equation}
\label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx) =  0 \quad \text{and} \quad
  \kappa(\vx, \vx')
  = \frac{1}{\delta} \Jac{\weights^*}{\vx} \, \JacT{\weights^*}{\vx'},
\end{equation}
where the kernel is the so-called Neural Tangent Kernel \citep[NTK,][]{jacot2018neural}.
We can combine this kernel function with the data set $\dataset$ to construct the posterior.
%
\citet{khan2019approximate} took a similar approach, but instead of fitting the GP posterior to the actual $\vy$, they rely on a transformation of $\vy$. Similarly, \citet{immer2021improving} obtain the same covariance function but use a different mean because they rely on a first-order approximation of the NN to form a Bayesian GLM model. Both approaches adjust the GP posterior mean function to ensure the predictions are from the NN $f_{\vw^*}(\vx)$.
In contrast, our method makes predictions directly from the GP and avoids such adjustments.
In the next section, we present our method for converting a trained NN to a sparse functional representation.
% on the non-sparse approximation.
% of our
% based on the non-sparse approximation.
% we will present the general formulation of the sparse functional representation of our trained neural network based on the non-sparse approximation.

% \newpage
\section{\our: Sparse function-space representation of neural networks}\label{sec:sfr}
%\paragraph{GP in the dual parameters}
In this section we present our method, named \our, which converts a trained NN into a GP.
\our is built upon a dual parameterization of the GP posterior, i.e. an alternative parameterization.
Early work by \citet{csato2002sparse} showed that Expectation Propagation gives rise to a dual paramaterization.
More recent work by \citet{adam2021dual,chang2023memory} showed a different dual parameterization arising from the evidence lower bound for sparse variational GPs \citep{hensman2013gaussian,titsias2009variational}.
The dual parameterization from \citet{adam2021dual,chang2023memory}, consists of parameters $\valpha$ and $\vbeta$, which gives rise to the predictive posterior,
% , we can build on the insight from \citet{csato2002sparse} to obtain a `dual' parameterization, $\valpha$ and $\vbeta$,
% The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
% %
\begin{align}  \label{eq:gp_pred}
  \myexpect_{p(f_i \mid\vy)}[f_i] &= \vk_{\vx i}^\top \valpha, \\
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i}, \nonumber
\end{align}
%
where $f_i = f_\vw(\vx_i)$ is the function output at input $\vx_i$,
the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.
\cref{eq:gp_pred} states that the first two moments of the resultant posterior process (which may not be a GP), can be parameterized via the dual
parameters $\valpha, \vbeta \in \R^{N}$,
% which are the vectors
defined as, \looseness-1
%
\begin{equation}
\begin{split}
  \label{eq:dual_param}
  \alpha_i &\coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f}\log p(y_i \mid f) |_{f=f_i}], \\
  %\quad \text{and} \quad
  \beta_i &\coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f f}\log p(y_i \mid f_i) |_{f=f_i}].
  \end{split}
\end{equation}
%
The relationships specified are valid for generic likelihoods, and involve no approximations since the expectation is under the exact posterior, but given that the model can be expressed in a kernel formulation. \cref{eq:gp_pred} and \cref{eq:dual_param} highlight that the approximate inference technique, usually viewed as a posterior approximation, can be alternatively interpreted as an approximation of the expectation of gradients of loss/likelihoods.

% Originally \citet{csato2002sparse} iteratively find dual parameters using an expectation propagation \citep[EP,][]{minka2001expectation} method. More recently, \citet{khan2017conjugate,adam2021dual} have shown this relationship for variational Gaussian processes where the above expectation is with respect to the approximate variational posterior. Meanwhile in \citet{wilkinson2023bayes}, they show links between linearization methods and how they solve the \cref{eq:dual_param}.

\paragraph{Dual parameters from NN}
Given that we use a Laplace approximation of the NN, we remove the expectation over the posterior \citep[see Ch.~3.4.1 in][for derivation]{rasmussen2006gaussian} and our dual parameters are given by,
%
\begin{align}
  \label{eq:dual_param_laplace}
  \hat{\alpha}_i &\coloneqq \nabla_{f}\log p(y_i \mid f) |_{f=f_i} , \\
  %\quad \text{and} \quad
  \hat{\beta}_i &\coloneqq - \nabla^2_{ff}\log p(y_i \mid f) |_{f=f_i}.
\end{align}
%
Substituting \cref{eq:dual_param_laplace} into \cref{eq:gp_pred}, we obtain our GP based on the converged NN. Again, this is similar to what was derived in \citet{immer2021improving} for the posterior variance function.
However, they use the NN's prediction $f_{\vw^*}(\cdot)$ for the posterior mean and do not emphasize the significance of the dual parameters.
The problem with the predictions in \cref{eq:gp_pred} is that they incur a cost of $O(N^3)$, which limits its application on large data sets.

\begin{table*}[t!]
  \centering\scriptsize
  \caption{Comparisons on UCI data with negative log predictive density (NLPD\textcolor{gray}{\footnotesize$\pm$std}, lower better). \our (with $M=20\% \text{ of } N$) is on par with the Laplace approximation (BNN/GLM) and outperforms the GP subset when the prior precision $(\delta)$ is tuned (right). Interestingly, when the prior precision ($\delta$) is not tuned (left), \our outperforms all other methods. The GP subset uses the same inducing points as \our.} %Results for methods marked with * as reported in the original benchmark~\citep{immer2021improving}.} %See \cref{app:uci} for additional tables with comparisons.}
	\label{tbl:uci}
	% \vspace*{-4pt}

	% Control table spacing
	% \renewcommand{\arraystretch}{1.}
	\setlength{\tabcolsep}{1.8pt}
	% \setlength{\tblw}{0.083\textwidth}

@@

\paragraph{Sparsification via dual parameters}
\label{sec:sparse-dual-gp}
%
Sparse GPs reduce the computational complexity by representing the GP as a low-rank approximation induced by a sparse set of input points \citep[see][for an early overview]{quinonero2005unifying}. Given that we have computed the dual parameters derived from our NN predictions and a kernel function, we could essentially employ any of these sparsification methods. In this work, we opt for the approach suggested by \citet{titsias2009variational} \citep[also used in the DTC approximation, see][]{quinonero2005unifying}, which defines the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) \, q(\vu) \, \mathrm{d}\vu$. Given that we have $p(f_i \mid \vu)$ determined by our GP prior, the goal is to find a $q(\vu)$.
